Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2016-01-25 cordillera-cycle.tex
Clarify grounded areas and s.l.-relevant volumes.

      Despite large differences in the input climate forcing
      (Fig.~\ref{fig:lr_ts}, upper panel), model output presents consistent
      features that can be observed across the range of forcing data used.
      In all simulations, modelled ice volumes remain relatively low during
      most of the glacial cycle, except during two major glacial events
      which occur between 62.2 and 56.9\,\unit{ka} during MIS~4, and between
      23.1 and 16.8\,\unit{ka} during MIS~2 (Fig.~\ref{fig:lr_ts}, lower
      panel). An ice volume minimum is consistently reached between 56.0 and
      42.9\,\unit{ka} during MIS~3. However, the magnitude and precise
      timing of these three events depend significantly on the choice of
      proxy record used to derive a~time-dependent climate forcing
      (Table~\ref{tab:extrema}).

      Simulations forced by the Greenland ice core palaeo-temperature
      records (GRIP, NGRIP) produce the highest variability in modelled ice
      volume throughout the last glacial cycle. In contrast, simulations
      driven by oceanic (ODP~1012, ODP~1020) and Antarctic (EPICA, Vostok)
      palaeo-temperature records generally result in lower ice volume
      variability throughout the simulation length, resulting in lower
      modelled ice volumes during MIS~4 and larger ice volumes during MIS~3.
      The NGRIP climate forcing is the only one that results in a~larger ice
      volume during MIS~4 than during MIS~2.

      While simulations driven by the GRIP and the two Antarctic
      palaeo-temperature records attain a~last ice volume maximum between
      19.1 and 16.8\,\unit{ka}, those informed by the NGRIP and the two
      oceanic palaeo-temperature records attain their maximum ice volumes
      thousands of years earlier. Moreover, the ODP~1012 run yields a~rapid
      deglaciation of the modelled area prior to 17\,\unit{ka}. The ODP~1020
      simulation predicts an early maximum in ice volume at 21.0\,\unit{ka},
      followed by slower deglaciation than modelled using the other
      palaeo-temperature records. Finally, whereas model runs forced by the
      Antarctic palaeo-temperature records result in a~rapid and
      uninterrupted deglaciation after the LGM, the simulation driven by the
      GRIP palaeo-temperature record also results in a~rapid deglaciation
      but in three steps, separated by two periods of ice sheet regrowth
      (Fig.~\ref{fig:lr_ts}).

@@

      Despite large differences in the timing of attained volume extrema
      (Table~\ref{tab:extrema}), all model runs show relatively consistent
      patterns of glaciation. During MIS~4, all simulations produce an
      extensive ice sheet, covering an area of at least half of that
      attained during MIS~2 (Table~\ref{tab:extrema};
      Fig.~\ref{fig:lr_maps}, upper panels). Corresponding maximum ice
      volumes also differ significantly between model runs, and vary between
      3.54 and~8.71\,\unit{m} sea level equivalents (m\,s.l.e.;
      Table~\ref{tab:extrema}).

      In the MIS~3 ice volume minimum reconstructions, a~central ice cap
      persists over the Skeena Mountains (Fig.~\ref{fig:lr_maps}, middle
      panels). Although this ice cap is present in all simulations, its
      dimensions depend sensitively on the choice of the applied
      palaeo-temperature record. Modelled ice volume minima spread over
      a~wide range between 1.54 and 2.94\,\unit{m}\,s.l.e.
      (Table~\ref{tab:extrema}).

      Modelled ice sheet geometries during the LGM (MIS~2;
      Fig.~\ref{fig:lr_maps}, lower panels) invariably include a~ca.
      1500\,\unit{km}-long central divide above 3000\,\unit{m\,a.s.l.}
      located along the spine of the Rocky Mountains. As an indirect result
      of the choice of scaling factors applied to different
      palaeo-temperature proxy records (Table~\ref{tab:records}), modelled
      maximum ice volumes also fall within a~tight range of 8.39 to
      9.07\,\unit{m}\,s.l.e. (Table~\ref{tab:extrema}).

@@

      Using the GRIP ice core palaeo-temperature record as climate forcing
      time-series, the model shows a~significant sensitivity to selected ice
      rheological and basal sliding parameters (Fig.~\ref{fig:sens_ts};
      Table~\ref{tab:sens_extrema}) in terms of modelled ice extent, and
      even more so in terms of modelled ice volumes.

      As a~direct result of the different scaling factors applied
      (Table~\ref{tab:sens_params}), the resulting simulations show very
      little difference in the modelled glaciated areas corresponding to
      maximum ice volumes during MIS~2, but also during MIS~4
      (Table~\ref{tab:sens_extrema}). However, such can not be said of the
      modelled glaciated area corresponding to minimum ice volume during
      MIS~3. In fact, the extent of the remnant ice cap which persists over
      the Skeena Mountains during this stage shows a~significant sensitivity
      to ice rheology of 31\,\unit{\%}, and an even more important
      sensitivity to basal sliding parameters of 62\,\unit{\%}
      (Table~\ref{tab:sens_extrema}).

      The modelled sea-level relevant ice volumes show more variability than
      the modelled glaciated areas (Table~\ref{tab:sens_extrema},
      Fig.~\ref{fig:sens_ts}). As one could expect, softer ice and weaker
      till both result in a~thinner ice sheet, while harder ice and stronger
      till result in a~thicker ice sheet. For instance, peak ice volume
      during the MIS~2 (LGM) varies by 30\,\unit{\%} between the two
      parametrisations of ice rheology used, and by 21\,\unit{\%} between
      the two parametrisations of basal sliding used. The differences in
      sea-level relevant ice volume are greatest during the MIS~3
      (Table~\ref{tab:sens_extrema}, Fig.~\ref{fig:sens_ts}) where both the
      areal and thickness contributions add up.

@@

      During MIS~2, the modelled total ice volume peaks at
      8.62\,\unit{m}\,s.l.e. (19.1\,\unit{ka}) in the GRIP simulation and at
      8.56\,\unit{m}\,s.l.e. (17.4\,\unit{ka}) in the EPICA simulation.
      However, these numers are subject to uncertainties on physics
      embedded in the model. The range of parameter values tested in our
      sensitivity study on the lower-resolution GRIP simulation
      (Sect.~\ref{sec:sens}) yielded relative errors of 30\,\unit{\%}
      regarding ice rheological parameters and 21\,\unit{\%} regarding basal
      sliding parameters (Fig.~\ref{fig:sens_ts};
      Table~\ref{tab:sens_extrema}).

@@

      During MIS~4, the modelled total ice volume peaks at
      7.43\,\unit{m}\,s.l.e. (57.6\,\unit{ka}) in the GRIP simulation and at
      4.84\,\unit{m}\,s.l.e. (61.9\,\unit{ka}) in the EPICA simulation,
      corresponding to respectively 86 and 57\,\unit{\%} of modelled MIS~2
      ice volumes.
      These estimates appear little sensitive to the range of parameter
      values tested in our sensitivity study (Sect.~\ref{sec:sens}), which
      yielded relative errors of 17\,\unit{\%} to ice rheological parameters
      and 21\,\unit{\%} to basal sliding parameters
      (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

      The modelled total ice volume corresponding to these persistent
      ice-dispersal centres attains a~minimum of 1.54\,\unit{m}\,s.l.e.
      (42.9\,\unit{ka}) in the GRIP simulation and of 2.55\,\unit{m}\,s.l.e.
      (52.4\,\unit{ka}) in the EPICA simulation, corresponding to
      respectively 18 and 30\,\unit{\%} of the MIS~2 ice volumes.
      However, these numbers should be considered with caution as our
      sensitivity study (Sect.~\ref{sec:sens}) shows that minimum ice volume
      during MIS~3 is highly sensitive to ice flow parameters, with relative
      errors of 57\,\unit{\%} to the range of ice rheological parameters
      tested and 120\,\unit{\%} to the range of basal sliding parameters
      tested (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

\begin{table*}
\caption{%
      Extremes in Cordilleran ice sheet volume and extent corresponding to
      MIS~4, 3 and 2 for each of the six low-resolution simulations
      (Fig.~\ref{fig:lr_ts}).}
\label{tab:extrema}
%\scalebox{.50}[.50]
{\begin{tabular}{l|ccc|ccc|ccc}
    \tophline
             & \multicolumn{3}{c|}{Age (ka)}
             & \multicolumn{3}{c|}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m\,s.l.e.)} \\
    Record   &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    GRIP     &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    NGRIP    &  60.27 &  45.87 &  22.85
             &   2.13 &   0.73 &   2.11
             &   8.71 &   1.70 &   8.60 \\
    EPICA    &  61.90 &  52.40 &  17.36
             &   1.48 &   0.98 &   2.08
             &   4.84 &   2.55 &   8.56 \\
    Vostok   &  62.17 &  55.95 &  16.83
             &   1.52 &   1.04 &   2.17
             &   5.04 &   2.94 &   9.07 \\
    ODP 1012 &  56.91 &  47.69 &  23.10
             &   1.34 &   0.87 &   2.06
             &   4.10 &   2.20 &   8.40 \\
    ODP 1020 &  60.72 &  52.88 &  20.99
             &   1.21 &   0.69 &   2.06
             &   3.54 &   1.54 &   8.39 \\
    \cline{1-10}
    Minimum  &  56.91 &  42.91 &  16.83
             &   1.21 &   0.67 &   2.06
             &   3.54 &   1.54 &   8.39 \\
    Maximum  &  62.17 &  55.95 &  23.10
             &   2.13 &   1.04 &   2.17
             &   8.71 &   2.94 &   9.07 \\
    \bottomhline
\end{tabular}}
\end{table*}

\begin{table*}[p]
\caption{%
      Extremes in Cordilleran ice sheet volume and extent corresponding to
      MIS~4, 3 and 2 using the GRIP paleo-climate forcing with each
      parameter configuration (Fig.~3). Relative differences (R. diff.) give
      rough error estimates related to varying selected ice rheology and
      basal sliding parameters (Table~\ref{tab:sens_params}).}
\label{tab:sens_extrema}
%\centering\makebox[\textwidth]
{\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c|}{Age (ka)}
             & \multicolumn{3}{c|}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m\,s.l.e.)} \\
    Config.  &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    Default  &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    \cline{1-10}
    Soft ice &  58.89 &  49.97 &  21.57
             &   1.96 &   0.54 &   2.08
             &   6.58 &   1.03 &   6.88 \\
    Hard ice &  57.32 &  42.90 &  19.14
             &   1.90 &   0.75 &   2.12
             &   7.83 &   1.91 &   9.46 \\
    R. diff. &    3\,\unit{\%} &   16\,\unit{\%} &   13\,\unit{\%}
             &    3\,\unit{\%} &   31\,\unit{\%} &    2\,\unit{\%}
             &   17\,\unit{\%} &   57\,\unit{\%} &   30\,\unit{\%} \\
    \cline{1-10}
    Soft bed &  58.90 &  49.21 &  19.53
             &   1.88 &   0.55 &   2.05
             &   6.46 &   1.03 &   7.52 \\
    Hard bed &  57.31 &  42.91 &  19.14
             &   1.93 &   0.96 &   2.13
             &   7.99 &   2.89 &   9.31 \\
    R. diff. &    3\,\unit{\%} &   15\,\unit{\%} &    2\,\unit{\%}
             &    3\,\unit{\%} &   62\,\unit{\%} &    4\,\unit{\%}
             &   21\,\unit{\%} &  120\,\unit{\%} &   21\,\unit{\%} \\
    \bottomhline
\end{tabular}}
\end{table*}

@@

\begin{figure*}%
\includegraphics{lr_ts}
\caption{%
      Temperature offset time-series from ice core and ocean records
      (Table~\ref{tab:records}) used as palaeo-climate forcing for the ice
      sheet model (top panel), and modelled ice volume (bottom panel)
      through the last 120\,\unit{ka}. Ice volumes are expressed in meters
      of sea level equivalent (\unit{m\,s.l.e.}). Gray fields indicate Marine
      Oxygen Isotope Stage (MIS) boundaries for MIS~2 and MIS~4 according to
      a~global compilation of benthic \chem{\delta^{18}O} records
      \citep{Lisiecki.Raymo.2005}. Hatched rectangles highlight the
      time-volume span for ice volume extremes corresponding to MIS~4
      (61.9--56.5\,\unit{ka}), MIS~3 (53.0--41.3\,\unit{ka}), and MIS~2
      (LGM, 23.2--16.8\,\unit{ka}). Dotted lines correspond to GRIP- and
      EPICA-driven 5\,\unit{km}-resolution runs.}
\label{fig:lr_ts}%
\end{figure*}%

\begin{figure*}%
\includegraphics{lr_maps}
\caption{%
      Snapshots of modelled surface topography (500\,\unit{m} contours)
      corresponding to the ice volume extremes indicated on
      Fig.~\ref{fig:lr_ts}. An ice cap persists over the Skeena Mountains
      (SM) during MIS~3. Note the occurence of spatial similarities despite
      large differences in timing.}
\label{fig:lr_maps}%
\end{figure*}%

@@

\begin{figure*}%
\includegraphics{hr_maps_mis2}
\caption{%
      Modelled surface topography (200\,\unit{m} contours) and surface
      velocity (colour mapping) corresponding to the maximum ice volume
      during MIS~2 in the GRIP and EPICA high-resolution simulations.}
\label{fig:hr_maps_mis2}%
\end{figure*}%

\begin{figure*}%
\includegraphics{hr_maps_mis4}
\caption{%
      Modelled surface topography (200\,\unit{m} contours) and surface
      velocity (colour mapping) corresponding to the maximum ice volume
      during MIS~4 in the GRIP and EPICA high-resolution simulations.}
\label{fig:hr_maps_mis4}%
\end{figure*}%

@@

\begin{figure}%
\includegraphics{hr_ts_deglac}
\caption{%
      Temperature offset time-series from the GRIP and EPICA ice core
      records (Table~\ref{tab:records}) (top panel), and modelled ice volume
      during the deglaciation, expressed in meters of sea-level equivalent
      (bottom panel).}
\label{fig:hr_ts_deglac}%
\end{figure}%

>>>>

      Despite large differences in the input climate forcing
      (Fig.~\ref{fig:lr_ts}, upper panel), model output presents consistent
      features that can be observed across the range of forcing data used.
      In all simulations, modelled sea-level relevant ice volumes remain
      relatively low during
      most of the glacial cycle, except during two major glacial events
      which occur between 62.2 and 56.9\,\unit{ka} during MIS~4, and between
      23.1 and 16.8\,\unit{ka} during MIS~2 (Fig.~\ref{fig:lr_ts}, lower
      panel). An sea-level relevant ice volume minimum is consistently
      reached between 56.0 and
      42.9\,\unit{ka} during MIS~3. However, the magnitude and precise
      timing of these three events depend significantly on the choice of
      proxy record used to derive a~time-dependent climate forcing
      (Table~\ref{tab:extrema}).

      Simulations forced by the Greenland ice core palaeo-temperature
      records (GRIP, NGRIP) produce the highest variability in modelled
      sea-level relevant ice
      volume throughout the last glacial cycle. In contrast, simulations
      driven by oceanic (ODP~1012, ODP~1020) and Antarctic (EPICA, Vostok)
      palaeo-temperature records generally result in lower ice volume
      variability throughout the simulation length, resulting in lower
      modelled ice volumes during MIS~4 and larger ice volumes during MIS~3.
      The NGRIP climate forcing is the only one that results in a~larger ice
      volume during MIS~4 than during MIS~2.

      While simulations driven by the GRIP and the two Antarctic
      palaeo-temperature records attain a~last sea-level relevant ice volume
      maximum between
      19.1 and 16.8\,\unit{ka}, those informed by the NGRIP and the two
      oceanic palaeo-temperature records attain their maximum ice volumes
      thousands of years earlier. Moreover, the ODP~1012 run yields a~rapid
      deglaciation of the modelled grounded ice extent prior to
      17\,\unit{ka}. The ODP~1020
      simulation predicts an early maximum in ice volume at 21.0\,\unit{ka},
      followed by slower deglaciation than modelled using the other
      palaeo-temperature records. Finally, whereas model runs forced by the
      Antarctic palaeo-temperature records result in a~rapid and
      uninterrupted deglaciation after the LGM, the simulation driven by the
      GRIP palaeo-temperature record also results in a~rapid deglaciation
      but in three steps, separated by two periods of ice sheet regrowth
      (Fig.~\ref{fig:lr_ts}).

@@

      Despite large differences in the timing of attained volume extrema
      (Table~\ref{tab:extrema}), all model runs show relatively consistent
      patterns of glaciation. During MIS~4, all simulations produce an
      extensive ice sheet, covering an area of at least half of that
      attained during MIS~2 (Table~\ref{tab:extrema};
      Fig.~\ref{fig:lr_maps}, upper panels). Corresponding maximum sea level
      potentials
      also differ significantly between model runs, and vary between
      3.54 and~8.71\,\unit{m} sea level equivalents (m\,s.l.e.;
      Table~\ref{tab:extrema}).

      In the MIS~3 ice volume minimum reconstructions, a~central ice cap
      persists over the Skeena Mountains (Fig.~\ref{fig:lr_maps}, middle
      panels). Although this ice cap is present in all simulations, its
      dimensions depend sensitively on the choice of the applied
      palaeo-temperature record. Modelled sea-level relevant ice volume
      minima spread over
      a~wide range between 1.54 and 2.94\,\unit{m}\,s.l.e.
      (Table~\ref{tab:extrema}).

      Modelled ice sheet geometries during the LGM (MIS~2;
      Fig.~\ref{fig:lr_maps}, lower panels) invariably include a~ca.
      1500\,\unit{km}-long central divide above 3000\,\unit{m\,a.s.l.}
      located along the spine of the Rocky Mountains. As an indirect result
      of the choice of scaling factors applied to different
      palaeo-temperature proxy records (Table~\ref{tab:records}), modelled
      maximum sea level potentials also fall within a~tight range of 8.39 to
      9.07\,\unit{m}\,s.l.e. (Table~\ref{tab:extrema}).

@@

      Using the GRIP ice core palaeo-temperature record as climate forcing
      time-series, the model shows a~significant sensitivity to selected ice
      rheological and basal sliding parameters (Fig.~\ref{fig:sens_ts};
      Table~\ref{tab:sens_extrema}) in terms of modelled grounded ice
      extent, and even more so in terms of modelled sea-level relevant ice
      volumes.

      As a~direct result of the different scaling factors applied
      (Table~\ref{tab:sens_params}), the resulting simulations show very
      little difference in the modelled glaciated areas corresponding to
      the maximum stage during MIS~2, but also during MIS~4
      (Table~\ref{tab:sens_extrema}). However, such can not be said of the
      modelled glaciated area corresponding to minimum stage during
      MIS~3. In fact, the extent of the remnant ice cap which persists over
      the Skeena Mountains during this stage shows a~significant sensitivity
      to ice rheology of 31\,\unit{\%}, and an even more important
      sensitivity to basal sliding parameters of 62\,\unit{\%}
      (Table~\ref{tab:sens_extrema}).

      The modelled sea-level potentials show more variability than
      the modelled glaciated areas (Table~\ref{tab:sens_extrema},
      Fig.~\ref{fig:sens_ts}). As one could expect, softer ice and weaker
      till both result in a~thinner ice sheet, while harder ice and stronger
      till result in a~thicker ice sheet. For instance, peak sea-level
      relevant ice volume
      during the MIS~2 (LGM) varies by 30\,\unit{\%} between the two
      parametrisations of ice rheology used, and by 21\,\unit{\%} between
      the two parametrisations of basal sliding used. The differences in
      ice volume are greatest during the MIS~3
      (Table~\ref{tab:sens_extrema}, Fig.~\ref{fig:sens_ts}) where both the
      areal and thickness contributions add up.

@@

      During MIS~2, the modelled sea level potential peaks at
      8.62\,\unit{m}\,s.l.e. (19.1\,\unit{ka}) in the GRIP simulation and at
      8.56\,\unit{m}\,s.l.e. (17.4\,\unit{ka}) in the EPICA simulation.
      However, these numers are subject to uncertainties on physics
      embedded in the model. The range of parameter values tested in our
      sensitivity study on the lower-resolution GRIP simulation
      (Sect.~\ref{sec:sens}) yielded relative errors of 30\,\unit{\%}
      regarding ice rheological parameters and 21\,\unit{\%} regarding basal
      sliding parameters (Fig.~\ref{fig:sens_ts};
      Table~\ref{tab:sens_extrema}).

@@

      During MIS~4, the modelled sea level potential peaks at
      7.43\,\unit{m}\,s.l.e. (57.6\,\unit{ka}) in the GRIP simulation and at
      4.84\,\unit{m}\,s.l.e. (61.9\,\unit{ka}) in the EPICA simulation,
      corresponding to respectively 86 and 57\,\unit{\%} of modelled MIS~2
      ice volumes.
      These estimates appear little sensitive to the range of parameter
      values tested in our sensitivity study (Sect.~\ref{sec:sens}), which
      yielded relative errors of 17\,\unit{\%} to ice rheological parameters
      and 21\,\unit{\%} to basal sliding parameters
      (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

      The modelled sea level potential corresponding to these persistent
      ice-dispersal centres attains a~minimum of 1.54\,\unit{m}\,s.l.e.
      (42.9\,\unit{ka}) in the GRIP simulation and of 2.55\,\unit{m}\,s.l.e.
      (52.4\,\unit{ka}) in the EPICA simulation, corresponding to
      respectively 18 and 30\,\unit{\%} of the MIS~2 ice volumes.
      However, these numbers should be considered with caution as our
      sensitivity study (Sect.~\ref{sec:sens}) shows that minimum ice volume
      during MIS~3 is highly sensitive to ice flow parameters, with relative
      errors of 57\,\unit{\%} to the range of ice rheological parameters
      tested and 120\,\unit{\%} to the range of basal sliding parameters
      tested (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

\begin{table*}
\caption{%
      Extremes in Cordilleran ice sheet grounded ice extent and sea-level
      relevant ice volume corresponding to
      MIS~4, 3 and 2 for each of the six low-resolution simulations
      (Fig.~\ref{fig:lr_ts}).}
\label{tab:extrema}
%\scalebox{.50}[.50]
{\begin{tabular}{l|ccc|ccc|ccc}
    \tophline
             & \multicolumn{3}{c|}{Age (ka)}
             & \multicolumn{3}{c|}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m\,s.l.e.)} \\
    Record   &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    GRIP     &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    NGRIP    &  60.27 &  45.87 &  22.85
             &   2.13 &   0.73 &   2.11
             &   8.71 &   1.70 &   8.60 \\
    EPICA    &  61.90 &  52.40 &  17.36
             &   1.48 &   0.98 &   2.08
             &   4.84 &   2.55 &   8.56 \\
    Vostok   &  62.17 &  55.95 &  16.83
             &   1.52 &   1.04 &   2.17
             &   5.04 &   2.94 &   9.07 \\
    ODP 1012 &  56.91 &  47.69 &  23.10
             &   1.34 &   0.87 &   2.06
             &   4.10 &   2.20 &   8.40 \\
    ODP 1020 &  60.72 &  52.88 &  20.99
             &   1.21 &   0.69 &   2.06
             &   3.54 &   1.54 &   8.39 \\
    \cline{1-10}
    Minimum  &  56.91 &  42.91 &  16.83
             &   1.21 &   0.67 &   2.06
             &   3.54 &   1.54 &   8.39 \\
    Maximum  &  62.17 &  55.95 &  23.10
             &   2.13 &   1.04 &   2.17
             &   8.71 &   2.94 &   9.07 \\
    \bottomhline
\end{tabular}}
\end{table*}

\begin{table*}[p]
\caption{%
      Extremes in Cordilleran ice sheet grounded ice extent and sea-level
      relevant ice volume corresponding to
      MIS~4, 3 and 2 using the GRIP paleo-climate forcing with each
      parameter configuration (Fig.~3). Relative differences (R. diff.) give
      rough error estimates related to varying selected ice rheology and
      basal sliding parameters (Table~\ref{tab:sens_params}).}
\label{tab:sens_extrema}
%\centering\makebox[\textwidth]
{\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c|}{Age (ka)}
             & \multicolumn{3}{c|}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m\,s.l.e.)} \\
    Config.  &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    Default  &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    \cline{1-10}
    Soft ice &  58.89 &  49.97 &  21.57
             &   1.96 &   0.54 &   2.08
             &   6.58 &   1.03 &   6.88 \\
    Hard ice &  57.32 &  42.90 &  19.14
             &   1.90 &   0.75 &   2.12
             &   7.83 &   1.91 &   9.46 \\
    R. diff. &    3\,\unit{\%} &   16\,\unit{\%} &   13\,\unit{\%}
             &    3\,\unit{\%} &   31\,\unit{\%} &    2\,\unit{\%}
             &   17\,\unit{\%} &   57\,\unit{\%} &   30\,\unit{\%} \\
    \cline{1-10}
    Soft bed &  58.90 &  49.21 &  19.53
             &   1.88 &   0.55 &   2.05
             &   6.46 &   1.03 &   7.52 \\
    Hard bed &  57.31 &  42.91 &  19.14
             &   1.93 &   0.96 &   2.13
             &   7.99 &   2.89 &   9.31 \\
    R. diff. &    3\,\unit{\%} &   15\,\unit{\%} &    2\,\unit{\%}
             &    3\,\unit{\%} &   62\,\unit{\%} &    4\,\unit{\%}
             &   21\,\unit{\%} &  120\,\unit{\%} &   21\,\unit{\%} \\
    \bottomhline
\end{tabular}}
\end{table*}

@@

\begin{figure*}%
\includegraphics{lr_ts}
\caption{%
      Temperature offset time-series from ice core and ocean records
      (Table~\ref{tab:records}) used as palaeo-climate forcing for the ice
      sheet model (top panel), and modelled sea-level relevant ice volume
      (bottom panel) through the last 120\,\unit{ka}, expressed in meters
      of sea level equivalent (\unit{m\,s.l.e.}). Gray fields indicate Marine
      Oxygen Isotope Stage (MIS) boundaries for MIS~2 and MIS~4 according to
      a~global compilation of benthic \chem{\delta^{18}O} records
      \citep{Lisiecki.Raymo.2005}. Hatched rectangles highlight the
      time-volume span for ice volume extremes corresponding to MIS~4
      (61.9--56.5\,\unit{ka}), MIS~3 (53.0--41.3\,\unit{ka}), and MIS~2
      (LGM, 23.2--16.8\,\unit{ka}). Dotted lines correspond to GRIP- and
      EPICA-driven 5\,\unit{km}-resolution runs.}
\label{fig:lr_ts}%
\end{figure*}%

\begin{figure*}%
\includegraphics{lr_maps}
\caption{%
      Snapshots of modelled surface topography (500\,\unit{m} contours)
      corresponding to the sea-level relevant ice volume extremes indicated
      on Fig.~\ref{fig:lr_ts}. An ice cap persists over the Skeena Mountains
      (SM) during MIS~3. Note the occurence of spatial similarities despite
      large differences in timing.}
\label{fig:lr_maps}%
\end{figure*}%

@@

\begin{figure*}%
\includegraphics{hr_maps_mis2}
\caption{%
      Modelled surface topography (200\,\unit{m} contours) and surface
      velocity (colour mapping) corresponding to the maximum ice
      sea-level relevant volume
      during MIS~2 in the GRIP and EPICA high-resolution simulations.}
\label{fig:hr_maps_mis2}%
\end{figure*}%

\begin{figure*}%
\includegraphics{hr_maps_mis4}
\caption{%
      Modelled surface topography (200\,\unit{m} contours) and surface
      velocity (colour mapping) corresponding to the maximum ice
      sea-level relevant volume
      during MIS~4 in the GRIP and EPICA high-resolution simulations.}
\label{fig:hr_maps_mis4}%
\end{figure*}%

@@

\begin{figure}%
\includegraphics{hr_ts_deglac}
\caption{%
      Temperature offset time-series from the GRIP and EPICA ice core
      records (Table~\ref{tab:records}) (top panel), and modelled sea-level
      relevant ice volume during the deglaciation (bottom panel).}
\label{fig:hr_ts_deglac}%
\end{figure}%
